Abstract Plots

### 2022-08-02 Abstract.Rmd
Reformatting of figure 2, replacing AUC for accuracy

# Diagram of Simulating Data
```{r, include=FALSE}
signature4.plot <- plot_96_profile(signature.4.mat)+
  labs(title = "Signature 4", x = "Context")+
  theme(plot.title = element_text(hjust = 0.5))

@@

# Scatterplot of Average AUC vs. Cosine Similarity to FFPE
```{r, include=FALSE}
set.seed(500)
#outer for loop: loop through each COSMIC v3 signature
average.auc.values <- c()
for (x in 1:length(signature.names)){
  signature.auc.vec <- c()
  #inner for loop: calculate misclassification rate 10x for each signature
  for (y in 1:10){
    #create a sample vector of 100 mutations for the COSMIC signature corresponding to number x
    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],500)
    #Create FFPE sample vector of 100 mutations
    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 100)
    classification.df.1 <- get_classification_df(list(sample.vec, ffpe.sample.vec), c(signature.names[x],"FFPE"), list(signature.matrix.list[[x]], ffpe.signature))
    class.df.colnames <- colnames(classification.df.1)
    sig.name <- class.df.colnames[3]
    classification.df.2 <- classification.df.1 %>%
      mutate(signature.indicator = ifelse(truth == sig.name, 1, 0))
    predictions <- classification.df.2[3]
    labels <- classification.df.2$signature.indicator
    pred <- prediction(predictions, labels)
    auc.perf <- performance(pred, measure = "auc")
    signature.auc.vec[y] <- auc.perf@y.values[[1]][1]
  }
  average.signature.auc <- mean(signature.auc.vec)
  average.auc.values[x] <- average.signature.auc
}
```

@@

<!-- ```{r} -->
<!-- step_load = seq(50, 1000, by = 50) -->
<!-- MMR.probabilities <- vector(mode = "list", length = length(step)) -->
<!-- MMR.truth <- vector(mode = "list", length = length(step)) -->

<!-- for (x in 1:length(step)){ -->
<!--   sample.ffpe <- create_signature_sample_vector(ffpe.sig, step[x]/2) -->
<!--   ffpe.truth <- rep("FFPE", step[x]/2) -->
<!--   sample.sig6 <- create_signature_sample_vector(cosmic.sig6, step[x]/2) -->
<!--   sig6.truth <- rep("MMR", step[x]/2) -->
<!--   test <- get_classification_df(list(sample.sig6, sample.ffpe), c("MMR","FFPE"), -->
<!--                                 list(cosmic.sig6, ffpe.sig)) -->
<!--   classification <- test %>% -->
<!--     mutate(trueBinary = ifelse(truth == "MMR",1,0)) -->

<!--   MMR.probabilities[[x]] <- classification$MMR -->
<!--   MMR.truth[[x]] <- classification$trueBinary -->
<!-- } -->
<!-- ``` -->

<!-- ```{r} -->
<!-- tob.probabilities <- vector(mode = "list", length = length(step)) -->
<!-- tob.truth <- vector(mode = "list", length = length(step)) -->

<!-- for (x in 1:length(step)){ -->
<!--   sample.ffpe <- create_signature_sample_vector(ffpe.sig, step[x]/2) -->
<!--   ffpe.truth <- rep("FFPE", step[x]/2) -->
<!--   sample.sig4 <- create_signature_sample_vector(cosmic.sig4, step[x]/2) -->
<!--   sig4.truth <- rep("tob", step[x]/2) -->
<!--   test <- get_classification_df(list(sample.sig4, sample.ffpe), c("Tobacco","FFPE"), -->
<!--                                 list(cosmic.sig4, ffpe.sig)) -->
<!--   classification <- test %>% -->
<!--     mutate(trueBinary = ifelse(truth == "Tobacco",1,0)) -->

<!--   tob.probabilities[[x]] <- classification$Tobacco -->
<!--   tob.truth[[x]] <- classification$trueBinary -->
<!-- } -->
<!-- ``` -->

<!-- ```{r} -->
<!-- load.MMR <- prediction(MMR.probabilities, MMR.truth) -->
<!-- load.tob <- prediction(tob.probabilities, tob.truth) -->
<!-- ``` -->

@@

<!-- ```{r} -->
<!-- load.sens.MMR.vec = evaluate_perf(100, length(step_load), load.MMR, "sens") -->
<!-- load.sens.tob.vec = evaluate_perf(100, length(step_load), load.tob, "sens") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- #plot(step, load.sens.MMR.vec, ylim = c(0,1)) -->
<!-- sensitivity = append(load.sens.MMR.vec, load.sens.tob.vec) -->
<!-- signature.sens = append(rep("MMR", length(load.sens.MMR.vec)), -->
<!--                         rep("Tobacco", length(load.sens.tob.vec))) -->
<!-- dilution = append(step, step) -->
<!-- sens.df <- data.frame(dilution, sensitivity, signature.sens) -->

<!-- ggplot(data = sens.df, mapping = aes(x = dilution,  -->
<!--                                      y = sensitivity,  -->
<!--                                      color = signature.sens)) + -->
<!--   geom_point() + -->
<!--   geom_smooth() + -->
<!--   labs(title = "Sensitivity as a function of mutational load", -->
<!--        x = "Number of mutations in each sample", -->
<!--        y = "Sensitivity value", -->
<!--        color = "True signal") + -->
<!--   coord_cartesian(ylim = c(0,1)) -->
<!-- ``` -->

@@

<!-- ```{r} -->
<!-- load.spec.MMR.vec = evaluate_perf(100, length(step.load), load.MMR, "spec") -->
<!-- load.spec.tob.vec = evaluate_perf(100, length(step.load), load.tob, "spec") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- #plot(step, load.spec.MMR.vec, ylim = c(0,1)) -->

<!-- specificity = append(load.spec.MMR.vec, load.spec.tob.vec) -->
<!-- signature.spec = append(rep("MMR", length(load.spec.MMR.vec)), -->
<!--                         rep("Tobacco", length(load.spec.tob.vec))) -->
<!-- dilution = append(step, step) -->
<!-- spec.df <- data.frame(dilution, specificity, signature.spec) -->

<!-- ggplot(data = spec.df, mapping = aes(x = dilution,  -->
<!--                                      y = specificity,  -->
<!--                                      color = signature.spec)) + -->
<!--   geom_point() + -->
<!--   geom_smooth() + -->
<!--   labs(title = "Specificity as a function of mutational load", -->
<!--        x = "Number of mutations in each sample", -->
<!--        y = "Specificity value", -->
<!--        color = "True signal") + -->
<!--   coord_cartesian(ylim = c(0,1)) -->
<!-- ``` -->

>>>>

# Diagram of Simulating Data

```{r, include=FALSE}
signature4.plot <- plot_96_profile(signature.4.mat)+
  labs(title = "Signature 4", x = "Context")+
  theme(plot.title = element_text(hjust = 0.5))

@@

# Accuracy vs Cosine similarity to FFPE

We first create a function that mixes mutations from FFPE and another given mutational signature and calls the bayes classifier `get_classification_df`

```{r}
test_mix_mutations <- function (cosmic_idx, num_cosmic, num_ffpe) {

  # Create baseline mutations
  baseline_sig <- signature.matrix.list[[cosmic_idx]]
  baseline_name <- signature.names[cosmic_idx]
  baseline_mut_vec <- create_signature_sample_vector(baseline_sig, num_cosmic)

  # Create FFPE mutations
  ffpe_mut_vec <- create_signature_sample_vector(ffpe.signature, num_ffpe)

  # Call excerno classification and format as a tibble
  get_classification_df (list(baseline_mut_vec, ffpe_mut_vec), 
                          c(baseline_name,"FFPE"), 
                          list(baseline_sig, ffpe.signature)) %>% 
    as_tibble()
}
```

And we try `test_mix_mutations` and plot the results to check it is working:

```{r}
set.seed(12345)
test_tbl <- test_mix_mutations(1,500,500)

test_tbl %>%
  select (mutations, truth, SBS1, FFPE) %>%
  pivot_longer(c(SBS1, FFPE)) %>%
  ggplot(aes(x=truth, y=value, fill=name)) +
    geom_boxplot()

Now let's create a function that calculates the performance metrics (sensitivity, specificity, accuracy) on the output of `test_mix_mutations`

```{r}
calc_metrics_mix <- function (cosmic_idx, num_cosmic, num_ffpe) {

  baseline_name <- signature.names[cosmic_idx]
  test_tbl <- test_mix_mutations(cosmic_idx,num_cosmic,num_ffpe)

  tp <- test_tbl %>%
    filter(truth == baseline_name & classify == baseline_name) %>%
    count() %>%
    pull(n)

  fp <- test_tbl %>%
    filter(truth != baseline_name & classify == baseline_name) %>%
    count() %>%
    pull(n)

  fn <- test_tbl %>%
    filter(truth == baseline_name & classify == "FFPE") %>%
    count() %>%
    pull(n)

  tn <- test_tbl %>%
    filter(truth == "FFPE" & classify == "FFPE") %>%
    count() %>%
    pull(n)

  tibble(
    Sensitivity = tp / (tp +fn),
    Specificity = tn / (tp + fn),
    Accuracy = (tp +tn) /(tp+tn+fp+fn))
}

And we test our method:

```{r}
set.seed(12345)

calc_metrics_mix(1,500,500)
calc_metrics_mix(4,500,500)
calc_metrics_mix(6,500,500)
```

We create a function that calculates the average metrics over a fixed number of simulations `num_iter` and test it

```{r}
calc_metrics_mix_num <- function (cosmic_idx, num_iter, num_cosmic, num_ffpe) {

  baseline_name <- signature.names[cosmic_idx]

  1:num_iter %>% 
    map_dfr(~calc_metrics_mix(cosmic_idx,num_cosmic,num_ffpe))  %>%
    pivot_longer(1:3) %>% 
    group_by(name) %>%
    summarize(mean = mean(value)) %>%
    mutate(sbs = baseline_name)

}

calc_metrics_mix_num(1,10,500,500)
calc_metrics_mix_num(4,10,500,500)
calc_metrics_mix_num(6,10,500,500)
```

Finally we are ready to generate the data over all of COSMIC mutational signatures, using 10 iterations and 500 mutations for FFPE and 500 mutations from the baseline COSMIC signature

```{r}
sbs_num = length(signature.names)
base_num = 500
ffpe_num = 500
reps = 10

set.seed(123456)
sim_tbl <- 1:sbs_num %>%
  map_dfr(~calc_metrics_mix_num(.x, reps, base_num, ffpe_num))
```

First we summarize the metrics across all the COSMIC mutational signatures

```{r}
library(kableExtra)

sim_tbl <- sim_tbl %>%
  mutate (name = factor (name, 
                         levels = c("Sensitivity", "Specificity","Accuracy"))) 
sim_tbl %>%
  ggplot(aes(mean))+
  geom_histogram(binwidth = 0.025)+
  facet_grid(.~name) +
  labs(x="Metric across all COSMIC signatures")

ggsave("paper_figures/supp3.png", dpi = 300, width = 175, height = 75, units = "mm")

summary_tbl <- sim_tbl %>%
  group_by(name) %>%
  summarize (median = median (mean),
           IQR = IQR(mean)) %>%
  mutate_if(is.numeric, format, digits=2)

summary_tbl %>%
  kbl() %>%
  kable_styling()
```

Now we explore the relationship of the performance metrics against the similarity between the baseline signature and FFPE using a figure

```{r}
library(broom)
sim_tbl <- sim_tbl %>%
  mutate (ffpe_cos = rep(ffpe.cosmic.cos.sims, each=3))

sim_tbl %>%
  ggplot(aes(x=ffpe_cos, mean)) +
  geom_point()+
  geom_smooth(method="lm", formula = "y~x")+
  facet_grid(~name)

And using a linear model

```{r}
lm_metric<- lm(mean ~ name:ffpe_cos, data = sim_tbl)

tidy(lm_metric) %>%
  kbl() %>%
  kable_styling()

glance(lm_metric) %>%
  kbl() %>%
  kable_styling()
```

We define a function to represent the mutational signatures that fall below a threshold for a particular metric.

```{r}
plot_metrics_excerno <- function (tbl, metric, threshold) {

  # Subset the table to the specific metric  
  new_tbl <- filter(tbl, name==metric) 
  # Identify signature that fall below the threshold
  outliers <- filter( new_tbl, mean < threshold)
  # Extract signatures SBS4 and SBS6 as a fixe reference in the graph
  fixed_sbs <- filter( new_tbl, sbs == "SBS4" | sbs == "SBS6")

  ggplot(new_tbl, aes(x = ffpe_cos, y = mean))+
    geom_point()+
    geom_smooth(method = "lm", se=FALSE)+
    geom_point(data = outliers, color = "red")+
    geom_point(data = fixed_sbs, color = "red")+
    geom_hline(yintercept=threshold, linetype="dashed", color = "red")+
    geom_label_repel(data=outliers, 
                   aes(label = sbs),
                   max.overlaps = 10, 
                   label.size = .1)+
    geom_label_repel(data=fixed_sbs, 
                   aes(label = sbs),
                   max.overlaps = 10, 
                   label.size =   0.1) +
   labs(x = "Cosine similarity against FFPE signature", 
         y = metric )
}
```

And we plot all the metrics using a threshold of 80%

```{r}
plot_metrics_excerno(sim_tbl, "Accuracy", 0.8)
ggsave("paper_figures/fig2b.png", dpi = 300, width = 175, height = 108, units = "mm")

plot_metrics_excerno(sim_tbl, "Sensitivity", 0.8)
ggsave("paper_figures/supp6.png", dpi = 300, width = 175, height = 108, units = "mm")

plot_metrics_excerno(sim_tbl, "Specificity", 0.8)
ggsave("paper_figures/supp7.png", dpi = 300, width = 175, height = 108, units = "mm")
```

And we obtain the linear model for each metric separately and print their information

```{r}
print_lm_coefs <- function (tbl, metric) {
  new_tbl <-  filter (tbl, name==metric)
  lm_metrics<- lm(mean ~ ffpe_cos, data = new_tbl)

  tidy(lm_metrics) %>%
    mutate_if(is.numeric, format, digits=2)%>%
    kbl() %>%
    kable_styling()
}

print_lm_fit <- function (tbl, metric) {
  new_tbl <-  filter (tbl, name==metric)
  lm_metrics<- lm(mean ~ ffpe_cos, data = new_tbl)

  glance(lm_metrics) %>%
    mutate_if(is.numeric, format, digits=2)%>%
    kbl() %>%
    kable_styling()
}

print_lm_coefs(sim_tbl, "Accuracy")
print_lm_fit(sim_tbl, "Accuracy")

print_lm_coefs(sim_tbl, "Sensitivity")
print_lm_fit(sim_tbl, "Sensitivity")

print_lm_coefs(sim_tbl, "Specificity")
print_lm_fit(sim_tbl, "Specificity")

# Scatterplot of Average AUC vs. Cosine Similarity to FFPE

```{r, include=FALSE}
set.seed(500)
#outer for loop: loop through each COSMIC v3 signature
average.auc.values <- c()
for (x in 1:length(signature.names)){
  signature.auc.vec <- c()
  #inner for loop: calculate misclassification rate 10x for each signature
  for (y in 1:10){
    #create a sample vector of 100 mutations for the COSMIC signature corresponding to number x
    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],500)
    #Create FFPE sample vector of 100 mutations
    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 500)
    classification.df.1 <- 
      get_classification_df(list(sample.vec, ffpe.sample.vec), 
                            c(signature.names[x],"FFPE"), 
                            list(signature.matrix.list[[x]], ffpe.signature))

    class.df.colnames <- colnames(classification.df.1)
    sig.name <- class.df.colnames[3]
    classification.df.2 <- classification.df.1 %>%
      mutate(signature.indicator = ifelse(truth == sig.name, 1, 0))
    predictions <- classification.df.2[3]
    labels <- classification.df.2$signature.indicator
    pred <- prediction(predictions, labels)
    auc.perf <- performance(pred, measure = "auc")
    signature.auc.vec[y] <- auc.perf@y.values[[1]][1]
  }
  average.signature.auc <- mean(signature.auc.vec)
  average.auc.values[x] <- average.signature.auc
}
```

@@

<!-- ```{r} -->

<!-- step_load = seq(50, 1000, by = 50) -->

<!-- MMR.probabilities <- vector(mode = "list", length = length(step)) -->

<!-- MMR.truth <- vector(mode = "list", length = length(step)) -->

<!-- for (x in 1:length(step)){ -->

<!--   sample.ffpe <- create_signature_sample_vector(ffpe.sig, step[x]/2) -->

<!--   ffpe.truth <- rep("FFPE", step[x]/2) -->

<!--   sample.sig6 <- create_signature_sample_vector(cosmic.sig6, step[x]/2) -->

<!--   sig6.truth <- rep("MMR", step[x]/2) -->

<!--   test <- get_classification_df(list(sample.sig6, sample.ffpe), c("MMR","FFPE"), -->

<!--                                 list(cosmic.sig6, ffpe.sig)) -->

<!--   classification <- test %>% -->

<!--     mutate(trueBinary = ifelse(truth == "MMR",1,0)) -->

<!--   MMR.probabilities[[x]] <- classification$MMR -->

<!--   MMR.truth[[x]] <- classification$trueBinary -->

<!-- } -->

<!-- ``` -->

<!-- ```{r} -->

<!-- tob.probabilities <- vector(mode = "list", length = length(step)) -->

<!-- tob.truth <- vector(mode = "list", length = length(step)) -->

<!-- for (x in 1:length(step)){ -->

<!--   sample.ffpe <- create_signature_sample_vector(ffpe.sig, step[x]/2) -->

<!--   ffpe.truth <- rep("FFPE", step[x]/2) -->

<!--   sample.sig4 <- create_signature_sample_vector(cosmic.sig4, step[x]/2) -->

<!--   sig4.truth <- rep("tob", step[x]/2) -->

<!--   test <- get_classification_df(list(sample.sig4, sample.ffpe), c("Tobacco","FFPE"), -->

<!--                                 list(cosmic.sig4, ffpe.sig)) -->

<!--   classification <- test %>% -->

<!--     mutate(trueBinary = ifelse(truth == "Tobacco",1,0)) -->

<!--   tob.probabilities[[x]] <- classification$Tobacco -->

<!--   tob.truth[[x]] <- classification$trueBinary -->

<!-- } -->

<!-- ``` -->

<!-- ```{r} -->

<!-- load.MMR <- prediction(MMR.probabilities, MMR.truth) -->

<!-- load.tob <- prediction(tob.probabilities, tob.truth) -->

<!-- ``` -->

@@

<!-- ```{r} -->

<!-- load.sens.MMR.vec = evaluate_perf(100, length(step_load), load.MMR, "sens") -->

<!-- load.sens.tob.vec = evaluate_perf(100, length(step_load), load.tob, "sens") -->

<!-- ``` -->

<!-- ```{r} -->

<!-- #plot(step, load.sens.MMR.vec, ylim = c(0,1)) -->

<!-- sensitivity = append(load.sens.MMR.vec, load.sens.tob.vec) -->

<!-- signature.sens = append(rep("MMR", length(load.sens.MMR.vec)), -->

<!--                         rep("Tobacco", length(load.sens.tob.vec))) -->

<!-- dilution = append(step, step) -->

<!-- sens.df <- data.frame(dilution, sensitivity, signature.sens) -->

<!-- ggplot(data = sens.df, mapping = aes(x = dilution,  -->

<!--                                      y = sensitivity,  -->

<!--                                      color = signature.sens)) + -->

<!--   geom_point() + -->

<!--   geom_smooth() + -->

<!--   labs(title = "Sensitivity as a function of mutational load", -->

<!--        x = "Number of mutations in each sample", -->

<!--        y = "Sensitivity value", -->

<!--        color = "True signal") + -->

<!--   coord_cartesian(ylim = c(0,1)) -->

<!-- ``` -->

@@

<!-- ```{r} -->

<!-- load.spec.MMR.vec = evaluate_perf(100, length(step.load), load.MMR, "spec") -->

<!-- load.spec.tob.vec = evaluate_perf(100, length(step.load), load.tob, "spec") -->

<!-- ``` -->

<!-- ```{r} -->

<!-- #plot(step, load.spec.MMR.vec, ylim = c(0,1)) -->

<!-- specificity = append(load.spec.MMR.vec, load.spec.tob.vec) -->

<!-- signature.spec = append(rep("MMR", length(load.spec.MMR.vec)), -->

<!--                         rep("Tobacco", length(load.spec.tob.vec))) -->

<!-- dilution = append(step, step) -->

<!-- spec.df <- data.frame(dilution, specificity, signature.spec) -->

<!-- ggplot(data = spec.df, mapping = aes(x = dilution,  -->

<!--                                      y = specificity,  -->

<!--                                      color = signature.spec)) + -->

<!--   geom_point() + -->

<!--   geom_smooth() + -->

<!--   labs(title = "Specificity as a function of mutational load", -->

<!--        x = "Number of mutations in each sample", -->

<!--        y = "Specificity value", -->

<!--        color = "True signal") + -->

<!--   coord_cartesian(ylim = c(0,1)) -->

<!-- ``` -->
